---
title: "Seminar 3"
author: "S. Natalia Alvarado Pachon"
date: "25 of September, 2015"
output: html_document
---
# Using RStudio and Markdown

## Description

This document is the first attempt to use **Markdown** and **R**. I will use the R Code constructed in the last session of the class as main input and will show the main codes used for the outputs.

## Using RStudio

RStudio has built-in Datasets that users can use to generate outputs and practice their skills.

To access the list of built-in lists in RStudio, the user must type: `data()`
To select a dataset, the user can use the code: `data("swiss")`

To inspect the dataset, the user can use the code: `?swiss`

For this excersice, we chose the dataset ***USArrests***

List of ***Number of Crimes per State***
```{r echo=FALSE}
knitr::kable(USArrests)
```

## Arrests in the US when convictions are related to murders

Given that the typology of the arrest can affect dramatically the number of incidents and the violence of the crime, we chose to focus only on Arrests related to Murders.

```{r}
data("USArrests")
```

A summary of the information contained in the Dataset gives us a better detail of it

```{r}

summary(USArrests$Murder)

```

The boxplot for this Dataset:

```{r}

boxplot(USArrests$Murder, main = '# of Murders per State')

```

summary(USArrests$Murder)
boxplot(USArrests$Murder, main = '# of Murders per State')

### Central Tendency and Variation

The mean for the data in the US Arrest is ***`r mean(USArrests$Murder)`***

The median is ***`r median(USArrests$Murder)`***

A histrogram shows more clearly the distribution of the data: 

```{r}
hist(USArrests$Murder)
```




## Analysis on viorent crimes
### Combining some more variables together

Another built-in dataset ***state*** contains information related to the 50 states. Here, I combined ***state.abb*** and ***state.x77*** to the original ***USArrests*** dataframe and created new one called ***UScombined***, which now has abbrebiations and some demographics such as population, income, etc.


```{r}
# abbrebiations
ABB <- c(state.abb)
# demographics (population, income, ect.) 
STATESFACT <- data.frame(state.x77) 

# new dataframe w/ demographics 
UScombined <- cbind(ABB, USArrests, STATESFACT) 
```

The table below shows the whole dataframe.

```{r echo=FALSE, results='asis'}
knitr::kable(UScombined)
```

 
### Muder rate vs life expectancy

The scatter plot below shows the relation between murder rate and life expectancy.

```{R}
plot(UScombined$Life.Exp, UScombined$Murder, 
     main = 'Murder rate vs life expectancy by states', 
     xlab = 'Life expectancy (1969-71)', ylab = 'Murder rate (per 100,000) (1973)'
     )
```

The correlation test indicates that there are negative correlation of `r cor(UScombined$Life.Exp, UScombined$Murder)`.
```{R}
cor.test(UScombined$Life.Exp, UScombined$Murder)
```

The result from a bivariate linear regression is shown below.
```{R}
lmr <- lm(UScombined$Murder~UScombined$Life.Exp)
summary(lmr)
```
