---
title: "Seminar 3"
author: "S. Natalia Alvarado Pachon"
date: "25 of September, 2015"
output: html_document
---
# Using RStudio and Markdown

## Description

This document is the first attempt to use **Markdown** and **R**. I will use the R Code constructed in the last session of the class as main input and will show the main codes used for the outputs.

## Using RStudio

RStudio has built-in Datasets that users can use to generate outputs and practice their skills.

To access the list of built-in lists in RStudio, the user must type: `data()`
To select a dataset, the user can use the code: `data("swiss")`

To inspect the dataset, the user can use the code: `?swiss`

For this excersice, we chose the dataset ***USArrests***

List of ***Number of Crimes per State***
```{r echo=FALSE}
knitr::kable(USArrests)
```

## Arrests in the US when convictions are related to murders

Given that the typology of the arrest can affect dramatically the number of incidents and the violence of the crime, we chose to focus only on Arrests related to Murders.

```{r}
data("USArrests")
```

A summary of the information contained in the Dataset gives us a better detail of it

```{r}

summary(USArrests$Murder)

```

The boxplot for this Dataset:

```{r}

boxplot(USArrests$Murder, main = '# of Murders per State')

```

summary(USArrests$Murder)
boxplot(USArrests$Murder, main = '# of Murders per State')

### Central Tendency and Variation

The mean for the data in the US Arrest is ***`r mean(USArrests$Murder)`***

The median is ***`r median(USArrests$Murder)`***

A histrogram shows more clearly the distribution of the data: 

```{r}
hist(USArrests$Murder)
```




You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
