---
title: "Seminor - week 3"
author: "Noriko Magara"
date: "25 September, 2015"
output: html_document
---


# Violent Crime Rates by US State

## Preset dataset

I used "**USArrests**" dataset that contains the number of crimes per 100,000 people in each state in **1973**.
The three types of crimes are: **Murder**, **Assault**, and **Rape**, as well as **Urban population**.

Below is the summary statistics of the dataset. 
The first command `data("USArrests")` loads the preset dataset.
The second command `summary(USArrests)` gives summary statistics as shown below.

```{r}
data("USArrests")
summary(USArrests)
```

## Combining with more vectors
I also used "**state**" dataset that contains informations related to the 50 states. I combined "**state.abb**" and "**state.x77**" to the **USArrests** dataframe and created new one called "**UScombined**" that now has abbrebiations, population, income, etc.


```{r}
# abbrebiations
ABB <- c(state.abb)
# demographics (population, income, ect.) 
STATESFACT <- data.frame(state.x77) 

# new dataframe w/ demographics 
UScombined <- cbind(ABB, USArrests, STATESFACT) 
```

The table below shows the whole dataframe.

```{R echo=FALSE, results='asis'}
knitr::kable(UScombined)
```

## Analysis on Viorent Crimes 
###Muder rate vs life expectancy

The scatter plot below shows the relation between murder rate and life expectancy.

```{R}
plot(UScombined$Life.Exp, UScombined$Murder, 
     main = 'Murder rate vs life expectancy by states', 
     xlab = 'Life expectancy (1969-71)', ylab = 'Murder rate (per 100,000) (1973)'
     )
```

The correlation test indicates that there are negative correlation of `r cor(UScombined$Life.Exp, UScombined$Murder)`.
```{R}
cor.test(UScombined$Life.Exp, UScombined$Murder)
```

The result from a bivariate linear regression is shown below.
```{R}
lmr <- lm(UScombined$Murder~UScombined$Life.Exp)
summary(lmr)
```
